MSDS 6372 Project 1 Objective 1

### 2020-02-09 Codebook/MSDS6372.Project.O1.Rmd

```{r libraries}
library(tidyverse)
library(mice)
library(skimr)
library(corrplot)
library(car)
library(ISLR)
library(ggplot2)
library(gridExtra)
library(SamplingStrata)
library(rbin)
library(leaps)
```

@@

@@

```{r drop irrelevant variables}
#Dropping logical irrelevant variables: "id", "name", "host_id", "host_name", "last_reiview", "latitude", "longitude", "neighborhood","availability_365"
nyc2 <- select(nycraw, -c("id", "name", "host_id", "host_name", "last_review", "latitude", "longitude", "neighbourhood","availability_365"))
head(nyc2)
```

#### Dependent Variable Check
- Checking on dependent variable range to make sure if there are zero's to remove. It would not be free to stay in NYC.
```{r}
nyc2 <- nyc2[!(nyc2$price==0),]
invisible(view(nyc2))
```

@@

```{r zero variable check}
#Results show no zero variance variables, leave in all
skim(nyc3)
```

@@

#### Checking for Multicollinearity
- Multicollinearity will weaken the model
  - number_of_reviews and reviews_per_month are correlated at 55%
    - Removing reviews_per_month

@@

@@

#### Examining VIFs
- The below results show us there is no need to remove any variables

```{r examining VIFs}
full.model<-lm(price~.,data=nyc4)  # . means all variable not mpg
vif(full.model)[,3]^2

alias(lm(price~.,data=nyc4))
```

@@

```{r linearity check}
#nyc4 %>% pairs() No color model
pairs(nyc4,col=nyc4$neighbourhood_group) #Color by neighborhood

par(mfrow=c(2,2))
plot(full.model)
```

@@

#### Examining VIFs of Log Price Variable

```{r exmaining log-linear VIFs}
log.depend.model<-lm(lprice~.,data=log.nyc)  # . means all variable not mpg
vif(log.depend.model)[,3]^2

alias(lm(lprice~.,data=log.nyc))
```

#### Reviewing Linearity with Independent and Logged Dependent (Price) Variable
- Curved relationships with the numeric variables
  - Could require a quadratic or logarithmic transformation

```{r linearity check of log-linear model}
pairs(log.nyc,col=log.nyc$neighbourhood_group)

par(mfrow=c(2,2))
plot(log.depend.model)
```

@@

```{r created logged independent variables}
log.indep.nyc <- log.nyc %>% mutate(lreviews=log(number_of_reviews))
log.indep.nyc <- log.indep.nyc %>% mutate(lnights=log(minimum_nights))
log.indep.nyc <- log.indep.nyc %>% mutate(llistings=log(calculated_host_listings_count))
invisible(log.indep.nyc)

log.indep.nyc <- select(log.indep.nyc, -c("minimum_nights", "number_of_reviews", "calculated_host_listings_count"))
invisible(log.indep.nyc)
```

#### Examining VIFs of Log-Log Model

```{r examining log-log VIFs}
log.indep.model<-lm(lprice~.,data=log.indep.nyc)  # . means all variable not mpg
vif(log.indep.model)[,3]^2

alias(lm(lprice~.,data=log.indep.nyc))
```

@@

```{r linearity check of log-log model}
pairs(log.indep.nyc,col=log.indep.nyc$neighbourhood_group) #Color by neighborhood

par(mfrow=c(2,2))
plot(log.indep.model)
```

#### Continuous Variable Manipulation
- Since we are seeing large clouds of data but no linear trend with logged and unlogged data, we are going to move forward with binning the data to see if it will assist us in determining if there is a relationship between the continuous variables and log price

```{r rbin}
nyc5 <- rbin_winsorize(nyc4, price, number_of_reviews, 50, winsor_rate = 0.05)
nyc5

nyc4 %>% keep(is.numeric) %>% pairs() #[Add color]

par(mfrow=c(2,2))
plot(full.model)
```

@@

nyc.bins$reviewsBin <- var.bin(nyc.bins$number_of_reviews, bins = 50)
nyc.bins$nightsBin <- var.bin(nyc.bins$minimum_nights, bins = 50)
nyc.bins$listBin <- var.bin(nyc.bins$calculated_host_listings_count, bins = 10)

nyc.bins <- select(nyc.bins,-c("minimum_nights", "number_of_reviews", "calculated_host_listings_count"))
invisible(nyc.bins)
```

@@

```{r linearity check of binned independent variables}
nyc.bin.model <-lm(price~.,data=nyc.bins)
#nyc.bins  %>% pairs() No color model
pairs(nyc.bins,col=nyc.bins$neighbourhood_group) #Color by neighborhood

par(mfrow=c(2,2))
plot(nyc.bin.model)
```

#### Explore potential correlation Neighborhood v Price
- We have to this moment not be able to surface linearity relationships between our numerican independent varaibles and our dependt variable
- Next we will check for correltaion of the categorical variables: room_type & neighbourhood_group
- Without removing the ouliers of price above 400 it is near impossible to see if there is a difference per neighborhood. We have removed those prices above 400 to see if the 
- We see a strong chance of correlation between Price and Neighbourhood Group

```{r}
nyc.categorical <- nyc4[!(nyc4$price > 400),]
nrow(nyc4)
nrow(nyc.categorical)
plot(nyc.categorical$neighbourhood_group, nyc.categorical$price, xlab = "Neighbourhood Group", ylab = "Price", title = "Price v Neighbourhood Group Correlation Check", col=c(7,32,52,82,107)) 
```

@@

```{r}
plot(nyc.categorical$room_type, nyc.categorical$price, xlab = "Room Type", ylab = "Price", title = "Price v Room Type Correlation Check", col=c(7,32,52)) 
```

#### Reviewing Linearity with Numeric Variables w/ Price > 400 Outliers Removed

```{r linearity check w Price greater than 400 Outliers Removed}
nyc.cat.model<-lm(price~.,data=nyc.categorical) 
#nyc4 %>% pairs() No color model
pairs(nyc.categorical,col=nyc.categorical$neighbourhood_group) #Color by neighborhood

par(mfrow=c(2,2))
plot(nyc.cat.model)
```

#### Reviewing Linearity with Numeric Variables w/ Log-Linear model Price > 400 Outliers Removed
- Still no obvious linearity

```{r linearity check w Log-Linear model Price greater than 400 Outliers Removed}
log.nyc.outliers <- log.nyc[!(log.nyc$lprice > 400),]

log.nyc.outliers.model<-lm(lprice~.,data=log.nyc.outliers) 
pairs(log.nyc.outliers, col=log.nyc.outliers$neighbourhood_group) #Color by neighborhood

par(mfrow=c(2,2))
plot(log.nyc.outliers.model)
```

#### Modeling
- We are not seeing any linear correlation between the dependent and independent numeric varaibles
- We are seeing a strong chance of linear correlation between the dependent and independent categorical variables
- We have surfaced the best residuals assumptions matched in a log-linear model
  - Due to this we are moving forward with modeling a log-linear model with singular variables as well as all interaction terms
  - This is to add complexity to our model, we have a low number of varaibles to select from
    - In adding this complexity we are tryign to surface any possible linear variable interations that may contribute to our model
    - If these are surfaced we will go back and use graphical means to verify the model's discovery

```{r full model check for p-value significance}
nyc.model = lm(lprice~neighbourhood_group + room_type + neighbourhood_group:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count, data=log.nyc.outliers)
summary(nyc.model)
```

@@

```{r forward selection model}
nyc.fwd = regsubsets(lprice~neighbourhood_group + room_type + neighbourhood_group:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count, data=log.nyc.outliers, method="forward",nvmax=20)
summary(nyc.fwd)$adjr2
summary(nyc.fwd)$rss
summary(nyc.fwd)$bic
```

```{r forward selection model 2}
nyc.fwd2 = regsubsets(lprice~neighbourhood_group + room_type + neighbourhood_group:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count, data=log.nyc.outliers, method="forward",nvmax=50)
summary(nyc.fwd2)$adjr2
summary(nyc.fwd2)$rss
summary(nyc.fwd2)$bic
```

```{r backward selection model}
nyc.bck = regsubsets(lprice~neighbourhood_group + room_type + neighbourhood_group:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count, data=log.nyc.outliers, method="backward",nvmax=20)
summary(nyc.bck)$adjr2
summary(nyc.bck)$rss
summary(nyc.bck)$bic
```

```{r exhaustive selection model}
nyc.exh = regsubsets(lprice~neighbourhood_group + room_type + neighbourhood_group:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count, data=log.nyc.outliers, method="exhaustive",nvmax=20)
summary(nyc.exh)$adjr2
summary(nyc.exh)$rss
summary(nyc.exh)$bic
```

```{r seqrep selection model}
nyc.seq = regsubsets(lprice~neighbourhood_group + room_type + neighbourhood_group:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count, data=log.nyc.outliers, method="seqrep",nvmax=20)
summary(nyc.exh)$adjr2
summary(nyc.exh)$rss
summary(nyc.exh)$bic
```

#### Assumptions Check
- Risduals are normally distributed
  - Log-linear model has closest to normally distrubuted residuals from last plot
  - Near normal distriution of residuals
  - Envoking Central Limit Theorum due to such a large sample size
- Constant variance
  - Near normal QQ-plot

```{r residuals, constant variance assumption}
par(mfrow=c(2,2))
plot(log.nyc.outliers.model)
```

- Independence
  - Assumed
- Multicollinearity
  - Confirmed with VIFs and pairs plot that there is no mjultcollinearity occuring

```{r multicollinearity assumption two}
vif(log.depend.model)[,3]^2
```

```{r multicollinearity assumption}
pairs(log.nyc.outliers, col=log.nyc.outliers$neighbourhood_group) #Color by neighborhood
```

- Removed all prices over 400 to help reduce data set and uncover any correlation that was present between the dependent varaibles and independent categorical variables

```{r outliers assumption}
#log.nyc.outliers <- log.nyc[!(log.nyc$lprice > 400),]
```

>>>>

```{r libraries}
library(tidyverse)
library(mice)
library(skimr)
library(corrplot)
library(car)
library(ISLR)
library(ggplot2)
library(gridExtra)
library(SamplingStrata)
library(rbin)
library(leaps)
library(dplyr)
library(ggplot2)
library(geosphere)
library(broom)
library(plyr)
library(devtools)
options(scipen=999)
```

@@

#### Determined that Times Square is a hot destination point. Used Lat long metrics to build a new metric with distance to Times Square

```{r tsquare_distance metric}
#Creating a new, tsquare_distance (distance to Times Square in miles)
for (i in 1:nrow(nycraw)) {
  n <- (distHaversine(c(nycraw$latitude[i], -nycraw$longitude[i]), c(40.7580, 73.9855)) / 1609.344)
  nycraw$tsquare_distance[i] <- n
}
head(nycraw)
```

@@

```{r drop irrelevant variables}
#Dropping logical irrelevant variables: "id", "name", "host_id", "host_name", "last_review", "latitude", "longitude", "neighborhood"
nyc2 <- select(nycraw, -c("id", "name", "host_id", "host_name", "last_review", "latitude", "longitude", "neighbourhood_group"))
head(nyc2)
```

#### Zero Value Variable Check
- Checking on dependent variable range to make sure if there are zero's to remove. It would not be free to stay in NYC.
- Checking on independent variables that have '0' as a value. Seeing that "availability_365" would mean the AirBnB isn't available at all, we are dropping those from our data set.
```{r zero value variable check}
nyc2 <- nyc2[!(nyc2$price==0),]
nyc2 <- nyc2[!(nyc2$availability_365==0),]
invisible(view(nyc2))
```

@@

#```{r zero variable check}
#Results show no zero variance variables, leave in all
#skim(nyc3)
#```

@@

### Numerical v Numerical Multicollinearity
- Multicollinearity will weaken the model
  - number_of_reviews and reviews_per_month are correlated at 55%
    - Removing reviews_per_month

@@

#### Changing Price variable range
```{r}
nyc4 <- filter(nyc4, price >= 25 & price <= 400)
```

@@

@@

```{r linearity check}
#nyc4 %>% pairs() No color model
pairs(nyc4,col=nyc4$neighbourhood) #Color by neighborhood
```

@@

#### Reviewing Linearity with Log-Linear model: Independent and Logged Dependent (Price) Variable
- Curved relationships with the numeric variables
  - Could require a quadratic or logarithmic transformation

```{r linearity check of log-linear model}
pairs(log.nyc,col=log.nyc$neighbourhood)
```

@@

```{r created logged independent variables}
log.indep.nyc <- log.nyc %>% mutate(lreviews=log(number_of_reviews))
log.indep.nyc <- log.indep.nyc %>% mutate(lnights=log(minimum_nights))
log.indep.nyc <- log.indep.nyc %>% mutate(llistings=log(calculated_host_listings_count))
log.indep.nyc <- log.indep.nyc %>% mutate(lavailablility=log(availability_365))
log.indep.nyc <- log.indep.nyc %>% mutate(ltsqr=log(tsquare_distance))
invisible(log.indep.nyc)

log.indep.nyc <- select(log.indep.nyc, -c("minimum_nights", "number_of_reviews", "calculated_host_listings_count", "availability_365", "tsquare_distance"))
# Checking for -inf logged results
invisible(log.indep.nyc)
# Drop -inf log reults in lavailability
log.indep.nyc<-log.indep.nyc[!(log.indep.nyc$lavailablility=="-Inf"),]
invisible(log.indep.nyc)
```

@@

```{r linearity check of log-log model}
pairs(log.indep.nyc,col=log.indep.nyc$neighbourhood) #Color by neighborhood
```

@@

nyc.bins$reviewsBin <- var.bin(nyc.bins$number_of_reviews, bins = 50)
nyc.bins$nightsBin <- var.bin(nyc.bins$minimum_nights, bins = 50)
nyc.bins$availBin <- var.bin(nyc.bins$availability_365, bins = 50)
nyc.bins$listBin <- var.bin(nyc.bins$calculated_host_listings_count, bins = 10)

nyc.bins <- select(nyc.bins,-c("minimum_nights", "number_of_reviews", "calculated_host_listings_count", "availability_365"))
invisible(nyc.bins)
```

@@

```{r linearity check of binned independent variables}
nyc.bin.model <-lm(price~.,data=nyc.bins)
#nyc.bins  %>% pairs() No color model
pairs(nyc.bins,col=nyc.bins$neighbourhood) #Color by neighborhood
```

#### Explore potential correlation Neighborhood v Price
- We have to this moment not be able to surface linearity relationships between our numerican independent varaibles and our dependentt variable
- Next we will check for correltaion of the categorical variables: room_type & neighbourhood_group
- We see a strong chance of correlation between Price and Neighbourhood Group

```{r}
plot(nyc4$neighbourhood, nyc4$price, xlab = "Neighbourhood", ylab = "Price", title = "Price v Neighbourhood Correlation Check", col=c(7,32,52,82,107)) 
```

- Narrowing down neighborhoods (randomly) to confirm neighbourhood is a significant categorical metrics to keep in MLR

```{r}
nycNeighborhood <- ddply(nyc4,.(neighbourhood), function(x) x[sample(nrow(x),1),])

plot(nycNeighborhood$neighbourhood, nycNeighborhood$price, xlab = "Neighbourhood", ylab = "Price", title = "Price v Neighbourhood Correlation Check") 
```

@@

```{r}
plot(nyc4$room_type, nyc4$price, xlab = "Room Type", ylab = "Price", title = "Price v Room Type Correlation Check", col=c(7,32,52)) 
```

#### Modeling
- We are not seeing any linear correlation between the dependent and independent numeric varaibles
  - We have tried: linear regression, log-linear regression transformation, log-log regression transformation, binning, and outlier drop.
- We are seeing a strong linear correlation between the dependent and independent categorical variables
- We have surfaced the best residuals assumptions matched in a log-linear model
  - Due to this we are moving forward with modeling a log-linear model with singular variables as well as all interaction terms
  - This is to add complexity to our model, we have a low number of varaibles to select from
    - In adding this complexity we are tryign to surface any possible linear variable interations that may contribute to our model
    - If these are surfaced we will go back and use graphical means to verify the model's discovery

```{r full model check}
nyc.model = lm(lprice~neighbourhood + room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + availability_365 + tsquare_distance, data=log.nyc)
summary(nyc.model)

#Surfacing only significant neighborhoods
nyc.model2 = tidy(nyc.model)
options(scipen = 999)
invisible(nyc.model2)
nyc.modeldf <- nyc.model2[nyc.model2$p.value < 0.05,]
invisible(view(nyc.modeldf))
```

```{r}
#would reducing the data set to just sig values help?
```

@@

```{r intuitive model}
#To be built and provided by Reagan, there should include about 4 or so intuitive models and we take the one with the best Adj rsquared
```

```{r forward selection model}
nyc.fwd = regsubsets(lprice~neighbourhood + room_type + neighbourhood:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + availability_365 + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count + availability_365:minimum_nights + availability_365:number_of_reviews + availability_365:calculated_host_listings_count + tsquare_distance:minimum_nights + tsquare_distance:number_of_reviews + tsquare_distance:calculated_host_listings_count + tsquare_distance:availability_365, method = "forward", data=log.nyc, nvmax=20)
summary(nyc.fwd)$adjr2

#predict.regsubsets =function (object , newdata ,id ,...){
#  form=as.formula (object$call [[2]])
#  mat=model.matrix(form ,newdata )
#  coefi=coef(object ,id=id)
#  xvars=names(coefi)
#  mat[,xvars]%*%coefi
#}

#testASE<-c()
#note my index is to 20 since that what I set it in regsubsets
#for (i in 1:20){
#  predictions<-predict.regsubsets(object=reg.fwd,newdata=test,id=i) 
#  testASE[i]<-mean((log(test$AvgWinnings)-predictions)^2)
#}

#par(mfrow=c(1,1))
#plot(1:20,testASE,type="l",xlab="# of predictors",ylab="test vs train ASE",ylim=c(0.2,0.8))

#index<-which(testASE==min(testASE))
#points(index,testASE[index],col="red",pch=10)
#rss<-summary(reg.fwd)$rss
#lines(1:20,rss/100,lty=3,col="blue")  #Dividing by 100 since ASE=RSS/sample size

#coef(reg.final,XXXX)

#final.model<-lm(log(AvgWinnings)~Greens+AvgPutts+Save,data=golf)
#summary(final.model)
```

```{r backward selection model}
nyc.bck = regsubsets(lprice~neighbourhood + room_type + neighbourhood:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + availability_365 + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count + availability_365:minimum_nights + availability_365:number_of_reviews + availability_365:calculated_host_listings_count + tsquare_distance:minimum_nights + tsquare_distance:number_of_reviews + tsquare_distance:calculated_host_listings_count + tsquare_distance:availability_365, method="backward", data=log.nyc, nvmax=20)
summary(nyc.bck)$adjr2
#summary(nyc.bck)$rss
#summary(nyc.bck)$bic
```

```{r seqrep selection model}
nyc.seq = regsubsets(lprice~neighbourhood + room_type + neighbourhood:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + availability_365 + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count + availability_365:minimum_nights + availability_365:number_of_reviews + availability_365:calculated_host_listings_count + tsquare_distance:minimum_nights + tsquare_distance:number_of_reviews + tsquare_distance:calculated_host_listings_count + tsquare_distance:availability_365, method="seqrep", data=log.nyc, nvmax=20)
summary(nyc.seq)$adjr2
#summary(nyc.seq)$rss
#summary(nyc.seq)$bic
```

```{r exhaustive selection model}
#Due to an exhaustive method being too large. We are commenting out this model and moving forward with selective methods that will run
#nyc.exh = regsubsets(lprice~neighbourhood + room_type + neighbourhood:room_type + minimum_nights + number_of_reviews + calculated_host_listings_count + availability_365 + minimum_nights:number_of_reviews + minimum_nights:calculated_host_listings_count + number_of_reviews:calculated_host_listings_count + availability_365:minimum_nights + availability_365:number_of_reviews + availability_365:calculated_host_listings_count + tsquare_distance:minimum_nights + tsquare_distance:number_of_reviews + tsquare_distance:calculated_host_listings_count + tsquare_distance:availability_365, data=log.nyc, nvmax=20)
#summary(nyc.exh)$adjr2
#summary(nyc.exh)$rss
#summary(nyc.exh)$bic
```

#### Assumptions Check on Intuitive Model
- Residuals near normally distributed but still skewed
- Envoking Central Limit Theorum due to such a large sample size
- Constant Variance
  - The QQ Plot shows an extreme deviation from normality. Even with the Central Limit Theorum we do not feel comforablte moving forward.

```{r Assumptions Check on Intuitive Model}
par(mfrow=c(2,2))
full.model<-lm(price~.,data=nyc4)
plot(full.model)
```

#### Assumptions Check on Log-Linear Intuitive Model
- Risduals are near normally distributed with them being slightly off
  - Envoking Central Limit Theorum due to such a large sample size
  - Passed
- Constant variance
  - The QQ-plot is showing much less departure from normality
  - Passed

```{r Assumptions Check on Log-Linear Intuitive Model}
par(mfrow=c(2,2))
log.depend.model<-lm(lprice~.,data=log.nyc)
plot(log.depend.model)
```

- Independence
  - Assumed
  - Passed
- Multicollinearity
  - We are seeing a high VIF for our distance to Times Square

```{r Multicollinearity Assumption on Log-Linear Intuitive Model}
vif(log.depend.model)[,3]^2
```
